Two Decades of Smalltalk VM Development

### 2018-08-17 paper.tex
Adding a brief Acknowledgements section for those that have done JIT work.
Oops.  A comment was eliding an important sentence about type inference in Slang.
Acknowledge Ryan's contributing the MIPSLE simulator. Fix some capitalisation.

\subsection{Initial JIT Development and New Back Ends}
When the JIT was first being developed and was at a state where bytecoded methods could actually be translated to machine code, development fell into a very productive and enjoyable pattern.  As a bytecode was encountered for which the template had yet to be written execution would stop in the debugger, which would create a skeleton implementation for the template method from the MessageNotUnderstood exception. The programmer would implement the body of the method and execution would continue.  Were the new template definition to be incorrect and cause a bug during JIT generation, rather than when the generated machine code was run, then the debugger could be used to wind execution back to the beginning of the JITting of the current method and resume execution, repeating the cycle until disassembly looked good, somewhat like lemming debugging above.  As a new method was encountered and JITted yet more templates would be encountered.  Similarly as the abstract machine instruction to concrete machine instruction mappings were added these could be implemented in the debugger, an experience repeated when new back ends were added. Much of the JIT was written in this interactive live style.  This style encourages tool creation because one can create, extend and polish tools in the middle of debugging sessions, enhancing them to make sense of ones current predicament.  Immediate feedback makes tool investment cheap and cheerful.

\subsection{Debugging New GC Algorithms and Lemming Debugging}

@@

%% Acknowledgments
%%\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
%  This material is based upon work supported by the
%  \grantsponsor{GS100000001}{National Science
%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%  conclusions or recommendations expressed in this material are those
%  of the author and do not necessarily reflect the views of the
%  National Science Foundation.
%\end{acks}

>>>>

\subsection{Initial JIT development and new back ends}
When the JIT was first being developed and was at a state where bytecoded methods could actually be translated to machine code, development fell into a very productive and enjoyable pattern.  As a bytecode was encountered for which the template had yet to be written execution would stop in the debugger, which would create a skeleton implementation for the template method from the MessageNotUnderstood exception. The programmer would implement the body of the method and execution would continue.  Were the new template definition to be incorrect and cause a bug during JIT generation, rather than when the generated machine code was run, then the debugger could be used to wind execution back to the beginning of the JITting of the current method and resume execution, repeating the cycle until disassembly looked good, somewhat like lemming debugging above.  As a new method was encountered and JITted yet more templates would be encountered.  Similarly as the abstract machine instruction to concrete machine instruction mappings were added these could be implemented in the debugger, an experience repeated when new back ends were added. Much of the JIT was written in this interactive live style.  This style encourages tool creation because one can create, extend and polish tools in the middle of debugging sessions, enhancing them to make sense of ones current predicament.  Immediate feedback makes tool investment cheap and cheerful.

\subsection{Debugging new GC algorithms and Lemming Debugging}

@@

%% Acknowledgments
%%\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
\section*{Acknowledgements}
The authors would like to thank Ryan Macnak who write the MIPSLE sprocessor simulator and MIPSLE back end.
%  This material is based upon work supported by the
%  \grantsponsor{GS100000001}{National Science
%    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%  conclusions or recommendations expressed in this material are those
%  of the author and do not necessarily reflect the views of the
%  National Science Foundation.
%\end{acks}
